|

### 2021-09-13 _main.Rmd
End of day:

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy in Population Health Sciences
degreedate: TBC
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.Rmd"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: For Brendan McHugh
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations

@@

<!--- May need small bit here on converting from one measurement to the other i.e.
mg/dL to mmol/L--->

### Lipid fractions

@@

$$TC \approx LDLc - HDLc - kTG$$
where $k$ is 0.20 if measurements are in milligrams per decilitre (_mg/dl_) and 0.45 if measured in millimole per litre (_mmol/l_).

@@

Elevated LDL-c in the bloodstream, a condition also known as hypercholesterolaemia or hyperlipidaemia,[@nelson2013] can lead to atherosclerosis,[@libby2019] the build-up of fatty deposits in the blood vessels. These deposits constrict blood flow and can lead to vascular complications. Alternatively, part of the deposit can detach from the artery walls, forming a clot that can lead to a heart attach or stroke.[@libby2019] Globally, the prevalence of elevated cholesterol was estimated by the World Health Organisation to be approximately 40%.

### Statins

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-Hydroxy-3-Methylglutaryl-Coenzyme-A (HMG-CoA) into mevalonic acid, by competitively binding with HMG-CoA Reductase (HMG-CoA-R). This conversion limits the rate at which cholesterol can be produced, enabling statins to effectively reduce the production of LDL cholesterol.

(ref:statin-mechanisam-cap) **Statin mechanism of action:** Overview of the

@@

```{r eval = FALSE, statin-mechanisam, echo = FALSE, results="asis", fig.cap='(ref:statin-mechanisam-cap)', out.width='100%', fig.scap='(ref:statin-mechanisam-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

@@

<!----------------------------------------------------------------------------->
### Triangulating across study designs

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. However, each method is limited by its own unique biases. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these inherent differences in study design, and as a result, in biases.[@lawlor2016a] If several sources of evidence are available and point towards identical conclusions about a exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, one expected to increased the effect of the exposure and the other to attenuate it.[@lawlor2016a] As such, triangulating these results can provides us a middle-ground between the competing directions of bias. A triangulation approach can also prove useful in a prospective manner, helping to design new studies that are at risk of different sources of bias to that already available from the published literature.

@@

_**McGuinness, L. A.**, and L Schmidt. (2020) medrxivr: Accessing and searching medRxiv and bioRxiv preprint data in R." Journal of Open Source Software 5.54 2651. DOI: [10.21105/joss.02651](https://doi.org/10.21105/joss.02651)_

A paper introducing the open-source preprint search tool described in Chapter \@ref(sys-rev-tools-heading). As is common for journal articles describing software, the paper is intentionally short providing only a broad overview of the tool while extensive documentation is available from the project website (see \@ref(sys-rev-tools-intro) for more details).

_Hennessy, E. A., Acabchuk, R., Arnold, P. A., Dunn, A. G., Foo, Y. Z., Johnson, B. T., Geange, S. R., Haddaway, N. R., Nakagawa, S., Mapanga, W., Mengersen, K., Page, M., Sánchez-Tójar, A. Welch, V.,  **McGuinness L. A.** (2020). Ensuring Prevention Science Research is Synthesis-Ready for Immediate and Lasting Scientific Impact. MetaArXiv. DOI: [10.31222/osf.io/ptg9j](https://doi.org/10.31222/osf.io/ptg9j)_

The experience of extracting data for the systematic review in Chapter \@ref(sys-rev-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials). [Under 2nd round of review at Prevention Science]

@@

```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-quote)', echo = FALSE}
"It is surely a great criticism of our profession that we have not organised a critical summary by speciality or sub-speciality, up-dated periodically, of all relevant RCTS."  
```

(ref:sys-rev-quote) --- Archie Cochrane, 2000 [@cochrane20001931]

@@

## To Do

-   Examine how many times a primary study was included in a review\
-   Look at whether searching preprints made any difference to my results
-   Could apply ROBIS to previous reviews to see how they stack-up? (This may be done as part of Georgia's summer project)

## Extra studies to include:

* Smeet et al. reference in the CPRD analysis
* Larsson - 10.1136/bmj.j5375 [@larsson2017b]
* Statins and the risk of dementia Jick

@@

-   Evidence map - show the distribution of the different studies population across the world.
-   Living systematic review approach - update weekly based on medRxiv

## Aims

@@

### Search strategy

We will systematically search electronic bibliographic databases to identify potentially relevant records. The search strategy used in each database will be developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE) terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy will include terms related to lipids, lipid modifying treatments, and dementia and its sub-types, and will be designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table 3.2 below and the full draft search strategies for each database are attached to this protocol. To ensure that the study design filters are not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as Line 7 NOT Line 13 in Table 3.2) will be screened. If any potentially relevant studies are identified, their titles and abstracts will be searched for key terms that can be incorporated into the filters to improve search sensitivity.

The following databases will be searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. We will also search clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials.

The abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) will be searched. Grey literature will also be searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses will be accessed using the Open Access Theses and Dissertations portal. We will also search bioRxiv and medRxiv, preprint repositories using a tool built as part of this thesis, to identify potentially relevant studies. Finally, the reference lists of included studies will be searched by hand while studies citing included studies will be examined using Google Scholar (forward and reverse citation searching).

@@

The full search strategy for the Medline search, which was designed first and subsequently translated into the syntax for other databases is presented in Appendix \@ref(appendix-search-strategy).

### Study selection

Records will be imported into Endnote and deduplicated using the method outlined in Bramer et al. (2016).[@bramer2016] Screening (both title/abstract and full-text) will be performed using a combination of Endnote and Rayyan, a web based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records will be performed by the primary author, with a random selection of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. If this demonstrates a significant level of erroneous exclusion by the primary author a larger proportion will be dual-screened. Full-text screening will also be completed in full by the primary author. A second reviewer will screen a random sample of included and excluded records, in addition to any records identified by the first reviewer as being difficult to assess against the inclusion criteria. Reasons for exclusion at this stage will be recorded. Disagreements occurring during either stage of the screening process will be resolved through discussion with a senior colleague. A PRIMSA flow diagram will be produced to document how records moved through the review.[@zotero-766]

@@

@@

Following guidance for a dose response meta-analysis, we picked midpoint for all studies where bands were provided (extremely common in the case of lipids)

This analysis was unusual, as multiple methods of intergration and analysis were required with data spannign many forms.

@@

@@

Risk of bias assessment was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomized controlled trials were assessed using the RoB2 tool,[@sterne2019], non-randomized studies of interventions were assessed using the ROBINS-I tool,[@sterne2016] and non-randomized studies of exposures were assessed using the ROBINS-E tool.[CITE]

I opted to use the ROBINS-E tool, despite it still being in development,[@morganr2020] in place of other existing published tools such as the Newcastle-Ottowa scale (NOS).[@wells2000newcastle] There are two primary reasons for this: firstly, current thinking in assessing internal validity has moved away from scales.

<!-- TODO Cite release version of this tool, and just use domain based approach -->

Secondly, using a domain based to assess the risk of bias in studies of exposure allowed for better comparison and consistency across the tool used for different study designs, as they are all domain based.

<!-- TODO This next bit is not true -->

At present, no risk of bias assessment tool for Mendelian randomization studies is available. Bias in these studies was assessed with the help of an expert panel drawn from my supervisors and other external invited researchers.[CITE] The risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation, as identified by a review of risk of bias assessments in systematic reviews of MR studies.[@mamluk2020] A copy of the is available in Appendix \@ref(appendix-mr-rob). Advance results from this review of bias assessment in MR studies were obtained from contact with the authors.

Results from the risk-of-bias assessment will be visualized using a paired forest/risk-of-bias blobbogram, created using the `robvis` tool. This tool was developed as an adjunct to this thesis to aid in creating standard risk-of-bias figures.[@mcguinness2019] A summary of the tool is provided in Appendix \@ref(appendix-robvis).

<!-- TODO Cite MR tool here from malmuk, see info from Francesca -->

Many reviews use reporting guidelines instead - this is bad![@costa2011]

## Studification

## Following up with authors

As part of the IPD analysis, relevant cohort studies were identified and approached. In the first instance, the corresponding author on the main man

### Statistical analysis methods

@@

Need formula here, and description of different methods and quarrels re fixed effect vs random effects meta-analyses.

This anal

Will also need description of dose-response meta-analysis here, with reference to forthcoming book Julian sent on.

### Visualisation of results

@@

A summary of this tool is contained in Appendix \@ref (), .

@@

Preprints are a valuable vidence source (see Introduction) but their inclusion in a systematic reivew

Additionally, I followed preprints up over time to investigate whether all identified preprints included in the review were subsequently published (in which case preprints give you a look into the future, and a systematic review update would capture these reports) or alternatively, if some preprints were not published, then preprints provide a different.

@@

@@

>>>>

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy in Population Health Sciences
degreedate: TBC
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.md"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: For Brendan McHugh
covidStatment: |
  `r paste(readLines("front-and-back-matter/_covidStatment.Rmd"), collapse = '\n  ')`
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations

@@

### Lipid fractions {#intro-lipid-fractions}

@@

\begin{equation}
  TC \approx LDLc + HDLc + kTG
  (\#eq:total-cholesterol-formula)
\end{equation}

where $k$ is 0.20 if measurements are in milligrams per decilitre (_mg/dl_) and 0.45 if measured in millimole per litre (_mmol/l_).

@@

Elevated LDL-c in the bloodstream, a condition also known as hypercholesterolaemia or hyperlipidaemia,[@nelson2013] can lead to atherosclerosis,[@libby2019] the build-up of fatty deposits in the blood vessels. These deposits constrict blood flow and can lead to vascular complications. Alternatively, part of the deposit can detach from the artery walls, forming a clot that can lead to a heart attach or stroke.[@libby2019] Globally, the prevalence of elevated cholesterol was estimated by the World Health Organization to be approximately 40%.

### Statins {#intro-statins}

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-hydroxy-3-methylglutaryl-coenzyme-A (HMG-CoA) into mevalonate, by competitively binding with HMG-CoA reductase (HMG-CoA-R). This conversion is a key rate-limiting step in the cholesterol biosynthesis pathway (see Figure \@ref(statin-mechanisam)), enabling statins to reduce effectively the production of LDL cholesterol.

(ref:statin-mechanisam-cap) Overivew of statins mechanism of action, inhibiting HMG-CoA reductase which controls the conversion of HMG-CoA to mevalonate, the rate-limiting step in cholesterol biosynthesis.

@@

```{r statin-mechanisam, echo = FALSE, results="asis",fig.pos = "H", fig.align = 'center', fig.cap='(ref:statin-mechanisam-cap)', out.width='50%', fig.scap='(ref:statin-mechanisam-scap)'}
knitr::include_graphics(file.path("figures/background/statinPath.png"))
```

Several statin treatments have been widely available for some time (see Table \@ref(tab:statinOverview-table)). Depending on the statin and dosage prescribed, the average reduction in LDL-c concentrations ranges from 15% with low-intensity regimen (e.g. ravastatin 5 mg/day) up to 60% with a high-intensity regimen (e.g. rosuvastatin 80 mg/day).[@collins2016a; @law2003] Statins also vary with regard to their lipophilicity (the extent to which they are lipid soluble), affecting their localisation within the body, with hydophilic statins being concentrated in the liver and lipophilic statins circualting more widely.[@schachter2005] This may create a divide in the pleiotropic affects of statins with differing lipophilicity, particularly given the ability of lipophilic statins to permeate the blood brain barrier.[@sierra2011]

<!----------------------------------------------------------------------------->
<!-- TODO Review the look and feel of this table -->

(ref:statinOverview-caption) Overview of commonly-prescribed statins, summarising their approval date (US), properties and lipid-lowering effect.

(ref:statinOverview-scaption) Overview of common statins

```{r statinOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->
### Triangulating across study designs {#intro-triangulation}

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. However, each method is limited by its own biases. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these inherent differences in study design, and as a result, in biases.[@lawlor2016a] If several sources of evidence are available and point towards identical conclusions about an exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, strengthen the effect of the exposure and the other to attenuate it.[@lawlor2016a] As such, triangulating these results can provides us a middle-ground between the competing directions of bias. A triangulation approach can also prove useful in a prospective manner, helping to design new studies that are at risk of different sources of bias to that already available from the published literature.[@munafo2018]

@@

<!-- TODO Ideally the two preprints for the CPRD and systematic review papers will go here -->

_**McGuinness, L. A.**, and L Schmidt. (2020) medrxivr: Accessing and searching medRxiv and bioRxiv preprint data in R. Journal of Open Source Software 5.54 2651. DOI: [10.21105/joss.02651](https://doi.org/10.21105/joss.02651)_

A paper introducing the open-source preprint search tool described in Chapter \@ref(sys-rev-tools-heading). As is common for journal articles describing software, the paper is intentionally short providing only a broad overview of the tool while extensive documentation is available from the project website (see Section \@ref(sys-rev-tools-intro) for more details).

_Hennessy, E. A., Acabchuk, R., Arnold, P. A., Dunn, A. G., Foo, Y. Z., Johnson, B. T., Geange, S. R., Haddaway, N. R., Nakagawa, S., Mapanga, W., Mengersen, K., Page, M., Sánchez-Tójar, A. Welch, V.,  **McGuinness L. A.** (2021). Ensuring Prevention Science Research is Synthesis-Ready for Immediate and Lasting Scientific Impact. Prevention Science . DOI: [10.1007/s11121-021-01279-8](https://doi.org/10.1007/s11121-021-01279-8)_

The experience of extracting data for the systematic review in Chapter \@ref(sys-rev-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials). <!-- TODO Need to create author declarations section -->

@@

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---
```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-quote)', echo = FALSE}
"It is surely a great criticism of our profession that we have not organised a critical summary by speciality or sub-speciality, up-dated periodically, of all relevant RCTS."  
```

(ref:sys-rev-quote) --- Archibald Cochrane, 2000 [@cochrane1979]

@@

<!-- ## To Do -->

<!-- -   TODO Examine how many times a primary study was included in a review -->
<!-- -   TODO Look at whether searching preprints made any difference to my results -->
<!-- -  TODO Could apply ROBIS to previous reviews to see how they stack-up? (This may be done as part of Georgia's summer project) -->

<!-- TODO Extra studies to include: -->

<!-- * Smeet et al. reference in the CPRD analysis -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->
<!-- * Statins and the risk of dementia Jick -->

@@

The importance of

I found that

The results of this review are used to guide the primary analysis presented in Chapter \@ref(cprd-analysis-heading), in addition to forming a key evidence source used in the triangulation exercise presented in Chapter \@ref(discussion-heading).

<!----------------------------------------------------------------------------->
## Introduction {#sys-rev-intro}

@@

### Protocol

A pre-specified protocol for this analysis was registered using the Open Science Framework, and is available for inspection.[@mcguinnessluke2020]

### Contributions

This systematic review was performed in conjunction with a team of secondary reviewer, all of whom . These included Alexandra MacAleenana, Athena Sheppard, Matthew Lee and Lena Schmidt.

<!-- TODO Double check this list at the end of the project. -->

### Search strategy and and study selection

I systematically search electronic bibliographic databases to identify potentially relevant records. The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy will include terms related to lipids, lipid modifying treatments, and dementia and its sub-types, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full draft search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

<!----------------------------------------------------------------------------->
(ref:searchOverview-caption) Summary of systematic search by topic. The full search strategy including all terms and the number of hits per term is included in Appendix \@ref(appendix-search-strategy).

(ref:searchOverview-scaption) searchOverview

```{r searchOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

To ensure that the study design filters are not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as "8 NOT 14" in Table \@ref(tab:searchOverview-table)) was screened. If any potentially relevant studies are identified, their titles and abstracts were searched for key terms that can be incorporated into the filters to improve search sensitivity.

<!-- TODO need to comment on this feed-back process, or remove -->

The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection (for the specific databases available thi via the University). We will also search clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials. As the contents of the Web of Science Core COllection can vary by institution, the specific database searched via this platform are listed in Appendix \@ref(appendix-wos-databases).

The abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's & Dementia) were searched by hand. <!-- TODO how were these searched? Date and other range -->

 Grey literature will also be searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses were accessed using the Open Access Theses and Dissertations portal. We will also search bioRxiv and medRxiv, preprint repositories using a tool built as part of this thesis, to identify potentially relevant studies. Finally, the reference lists of included studies were searched by hand while studies citing included studies will be examined using Google Scholar (forward and reverse citation searching).

<!-- TODO Look at citationchaser for this, but have a cut-off publication date of July 2019 - better than searching by hand/Google Scholar -->

@@

#### Study selection

Records were imported into Endnote and deduplicated using the method outlined in Bramer et al. (2016).[@bramer2016] In summary, this method uses multiple stages to identify potential duplicates, beginning with automatic deletion of records matching on multiple fields ("Author" + "Year" + "Title" + "Journal"), followed by manual review of less similar articles (e.g. those matched based on the "Title" field alone).

Screening (both title/abstract and full-text) was performed using a combination of Endnote, a citation management tool,[@hupe2019] and Rayyan, a web-based screening application.[@ouzzani2016] Title and abstract screening to remove obviously irrelevant records was performed primarily by the primary author, with a random selection of excluded records being screened in duplicate to ensure consistency with the inclusion criteria.

Full-text screening will also be completed in full by the primary author. A second reviewer screened a random sample (approx 10%) of included and excluded records, in addition to any records identified by the first reviewer as being difficult to assess against the inclusion criteria. Reasons for exclusion at this stage were recorded. Disagreements occurring during either stage of the screening process were resolved through discussion with a senior colleague. A PRIMSA flow diagram will be produced to document how records moved through the review.[@page2021]

@@

<!-- TODO Expand on rationale for exclusion of cross-sectional studies. Dementia is a disease severly at risk of differential recall, leading to misclassification bias. By using prospective studies, the potential for this bias is reduced. -->

#### Assessment of inter-rater reliability

If this approach demonstrates a significant level of erroneous exclusion by the primary author a larger proportion will be dual-screened.

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure of inter-rater reliability was chosen over other methods of assessing inter-rater reliability such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review, this is an important characteristic.

Gwet's AC1 is defined as:

$$AC1 = \frac{observed\;agreement-chance\;agreement}{1-chance\;agreement}$$

In reference to a two-by-two table with cells A, B, C and D, it is calculated using the following:

\begin{equation}
  AC1 = \frac{\frac{A+D}{N}-e(\gamma)}{1-e(\gamma)}
  (\#eq:AC1-main)
\end{equation}

\begin{equation}
  q = \frac{(A+C)+(A+B)}{2N}
  (\#eq:AC1-supp)
\end{equation}

<!-- TODO Need to be sure of how to calculate. -->

How to interpret agreement co-efficients is widely debated, and while arbitary cut-off values may mislead,[@brennan1992] they provide a useful rubric by which to assess inter-rater agreement. Here, I used guidelines based on a stricter interpretation of the Cohen's $kappa$ coefficient,[@mchugh2012] presented in Table \@ref(tab:gwet-table).

@@

Following guidance for a dose response meta-analysis, I picked midpoint for all studies where bands were provided (extremely common in the case of lipids)

This analysis was unusual, as multiple methods of integration and analysis were required with data spanning many forms.

A piloted data extraction for was used. I extracted the data in the first instance and subset was checked by a colleague for accuracy.

@@

Where necessary, lipid levels reported in _mmol/L_ were converted in _mg/dL_ using the following formula:

\begin{equation} 
  mg/dL = mmol/L \times{} Z
  (\#eq:lipidConversion)
\end{equation}

where $Z = 38.67$ for total cholesterol, LDL-c and HDL-c, and $Z = 88.57$ for triglycerides. For widely-used categorises of lipids levels on the _mg/dL_ scale, see Table \@ref(tab:lipidLevels-table) in Section \@ref(intro-lipid-fractions).

#### Studification

Multiple publications resulting from the analysis of the same data were included and grouped ().

This was particularly relevant to preprints and published papers reporting the same study, which were not considered duplicate records. Instead, they were considered as different reports of the same study. This is due to the potential for the published version to offer some information that the preprint did not, and vice versa.

#### Following up with authors {#contacting-authors}

Where data points required either for the analysis or risk-of-bias assessment but were not reported, the primary authors of the study were contacted. This approach was taken due to the potentially large impact of following up with authors on the results of the review,[@reynders2019]

This was particular important for the dose response meta-analysis, we

Use the example of getting the cut-off points from the

#### Converting between different

WHile the analysis of change scores is feasible, as they were explicityl

@@

A key use of the review presented in this chapter is to identify different sources of evidence at risk of a diverse range of biases, and to contrast and compare findings across them (see Section \@ref(intro-triangulation) and Section \@ref(triangulation-overview) for an overview of triangulation). To enable this triangulation exercise, a detailed and structured risk of bias assessment formed an important part of this review.

There has been a recent movement within the evidence synthesis community from examining _methodological quality_ to assessing _risk of bias_,[@mcguinness2018; @sterne2016] and thus directly evaluating the internal validity of a study. Internal validity is defined here as the absence of systematic error (or bias) in a study, which may influence its results.[@campbell1957; @juni2001]

This move was prompted by a unclear definition of "methodological quality" which could include facets such as, and challenges in the comparison of results from different tools.

As part of this, the community also moved from checklist or score based tools towards domain-based methods, in which different potential sources of bias in a study are assessed in order.

Additionally, results should be assessed at the result (defined as a a specific outcome at a specific timepoint) rather than the study level. For example, a study may report on the efficacy of an intervention at 6 months and two years follow-up. missing outcome data that is not an issues at 6 months may introduce bias at 2 year follow-up. In this case, assigning a bias judgement to the study as a whole hides this differential biases for each result.

In this review, domain-based tools were used to assess the risk of bias for each result in each included study. The study design-specific tools are introduced and discussed in more detail in the following sections.

In addition, the tool also aim to capture the potential direction of bias for each given domain. An optional question at the end of each domain, e.g. for Domain 1 of the RoB2 tool: "What is the predicted direction of bias arising from the randomization process?", possible responses included

* Not applicable 
* Favours experimental 
* Favours comparator
* Towards null 
* Away from null 
* Unpredictable

#### Randomised controlled trials

Risk of bias assessment in randomised controlled trials was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomized controlled trials were assessed using the RoB2 tool,[@sterne2019], 
The tool assess the risk of bias across five domains: Bias arising from the randomization process, Bias due to deviations from intended intervention, Bias due to missing outcome data, Bias in measurement of the outcome, Bias in selection of the reported result. Acceptable judgements include: low risk of bias, some concerns, high risk of bias. Each of the 5 domains contains a series of signalling questions or prompts, which guide the user through the tool.

One a domain-level judgement for each domain has been assigned, an overall judgement, using the same three levels of risk of bias, is assigned to

#### Non-randomised studies of interventions/exposures

For non-randomised studies of interventions (NRSI), I used the ROBINS-I (Risk of bias in non-randomised studies - of interventions) <!-- TODO Capitalise appropriately --> tool.[@sterne2016] This tool assess the risk of bias across seven domains: Bias due to confounding, Bias due to selection of participants, Bias in classification of interventions, Bias due to deviations from intended interventions, Bias due to missing data, Bias in measurement of outcomes, and Bias in selection of the reported result. Similar to the RoB 2, it has a number of prompting questions per domain, with judgements including. Ideally, obs studies assessed in reference to an idealised randomised controlled trial. Further, the rare overall judgement of Low should be considered equivalent to a randomised controlled trial.

For non-randomised studies of exposures (NRSE), I opted to use a preliminary version of the ROBINS-E tool, despite it still being in development.[@morganr2020] While the signalling questions for each domain are yet to be confirmed, the overarching domains have been finalised.

Despite these limitations, I chose this approach in place of other existing published tools such as the Newcastle-Ottowa scale (NOS).[@wells2000] The motivation for this decision was two-fold. In the first instance, as mentioned in the introduction to this section, using a domain-based tool has distinct advantages over better-developed checklist-type tools such as the NOS. In addition, using a domain-based tool enabled better comparison with risk-of-bias assessments performed for the other study designs.

#### Mendelian randomisation studies

<!-- TODO This next bit is not all true -->

At present, no risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk of bias assessments in systematic reviews of MR studies. A copy of this tool is available in Appendix \@ref(appendix-mr-rob). Advance results from this review of bias assessment in MR studies were obtained from contact with the authors.

#### Risk of bias due to missing evidence

A recent shift towards the assessment of missing evidence due to selective non-reporting (as distinct from the selective reporting of a single result from multiple planned - handled in Domain 5 of the RoB 2 tool for example).

As part of this review, I piloted the RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool. <!-- TODO CITATION NEEDED --> Results were visualised

### Analysis methods

Studies were not combined across different study designs (i.e. RCTs were not combined in a meta-analysis with results from observational studies).

#### Standard meta-analysis

@@

Both a fixed-effect and random-effects meta-analysis model was.

For

Need formula here, and description of different methods and quarrels regarding distinction fixed effect vs random effects meta-analyses.

This analytical approach was used for

$u$

\begin{equation}
  \theta_i = \mu + u_i
  (\#eq:meta-analysis-random)
\end{equation}

\begin{equation}
  weighted\ average = \frac{\sum Y_i (1/SE_i^2)}{\sum(1/SE_i^2)}
  (\#eq:meta-analysis-fixed)
\end{equation}

where ...

<!-- TODO Need equation for random effects too. See Sean's thesis here. -->

This analytical approach was used for cases where the exposure was dichotomous (e.g. statin use or not, hypercholesterolemia or not).

#### Dose-response analyses

<!-- TODO Explore how to map from simple diagnosis of hyperchol to dose response. Seems that does response needs >3 categories -->
<!-- Will also need description of dose-response meta-analysis here, with reference to forthcoming book Julian sent on.  -->

Several of the included studies presented data on multiple categories of lipid levels, but provided an overall effect estimate based on a comparison of only two of these categories (e.g. for example, highest vs lowest quartile). <!-- TODO Question: if I look at highest vs lowest, and also at top dose vs lowest dose, am I double counting results? -->
 while this allows for easy interpretation of the resulting effect estimate, it ignores any potential non-linear relationships between the exposure and outcome, in addition to discarding useful information contain in the interim groups.

Where possible, data for all exposure groups was extracted, and this <!-- TODO ??? -->
. Studies were excluded from this analysis if the number of categories was less than three, if the exposure cut-off points for for each category were not presented (e.g. if the study reports splitting participants into quartiles and comparing the highest vs lowest without giving the quartile bands in mmol/L).

<!-- Taken from another paper: "A potential non-linear association was evaluated using a restricted cubic spline model with three knots at the 10th, 50th and 90th percentile of frequency of the exposure [@orsini2012meta]" -->

<!-- TODO Do I initially want to fit a linear relationship across categories and  -->
A XXXX spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of an exposure can have an effect versus the "normal" reference dose. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fractions, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

Due to the requirements for the dose response analysis, studies were excluded from this secondary analysis if they did not provide the require information: cut-off points. Where this was not reported in the study, I contacted the corresponding author to attempt to obtain the required information (see \@ref(contacting-authors)).

When the highest category was open ended (e.g LDL-c $\geqslant$ 200 mg/dL), I calculated category midpoint by assuming the widht of the higest category was the same as the one immediately below it. Similarly, when, the lowest category was open-ended (e.g LDL-c $\leqslant$ 100 mg/dL), I set the lower boundary for this category to zero (though this is unlikely to occur natually, it was difficult to define).

I assumed the width of the category to be the same as the adjacent category.
When the lowest category was open ended, I set the lower boundary to zero.

#### Sensitivity analyses

I conducted a leave one out analysis in order to explore the impact of any given study on the results. <!-- TODO CITATION NEEDED -->

In addition,

&nbsp;
<!----------------------------------------------------------------------------->

#### Visualisation of results

@@

This tool grew out of a collaboration with other reserachers to design a modular method for creating these plots <!-- Apparently collaboration is a big thing they want to see -->

A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this analysis were created using

Results from the risk-of-bias assessment were visualized using a paired forest/risk-of-bias blobbogram, created using the `robvis` tool. This tool was developed as an adjunct to this thesis to aid in creating standard risk-of-bias figures.[@mcguinness2020robvisPaper] A summary of the tool is provided in Appendix \@ref(appendix-robvis).

Also, the use of evidence maps is becoming increasingly common to explore the distibution of research cohorts included in a

As such, each individual cohort contibuting to the evidence base was quantified and visualised on a world map.

&nbsp;
<!----------------------------------------------------------------------------->

@@

Preprints are a valuable evidence source (see Introduction) but their inclusion in a systematic reivew
s
Additionally, I followed preprints up over time to investigate whether all identified preprints included in the review were subsequently published (in which case preprints give you a look into the future, and a systematic review update would capture these reports) or alternatively, if some preprints were not published, then preprints provide a different.

@@

Patient and public members were intended to be involved at this poit

### Deviations from the protocol

A dose response meta-analysis was not planned, marking a departure from the registered protocol. <!-- TODO Check this is accurate -->

Similarly, there was initially no intention to try and examine the risk of bias due to missing evidence.

@@

### Search and screening results

Following the search

`r num_to_text(23447,T)` records were identified through database searches. Following depulication, `r num_to_text(16109)` unique records remained.

No eligible records were identified in the screening of the 500 records excluded by the search. Many of these were basic science studies <!-- TODO text -->

```{r prisma-flow-setup, include = FALSE}
```

\blandscape{}

(ref:prisma-flow-cap) PRISMA Flow diagram illustrating how records moved through the systematic review process. The different contributions of databases and preprint servers to the review are indicated.

(ref:prisma-flow-scap) PRISMA flow diagram

```{r prisma-flow-fig, echo = FALSE, results="asis", fig.pos="H", fig.cap='(ref:prisma-flow-cap)', out.width='100%', fig.scap='(ref:prisma-flow-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

\elandscape{}

Following de-duplication, the titles and abstracts of `r num_to_text(16109)` records were assessed for eligibility. `r num_to_text(387)` were deemed potentially eligible and the full text records for these were requested and screened. \@ref(fig:prisma-flow-fig) <!-- TODO Check these figures -->
